---
title: 'Case Study: A Sales Report for Global Superstore'
author: "Daniel Kuhman"
date: "`r Sys.Date()`"
output: html_document
---

## Business Prompt
Welcome to Global Superstore. We’re an online retailer specializing in office items, including basic supplies, furniture, and tech products. We serve corporate, consumer, and home-based buyers. As our name suggests, we operate globally, with customers from 165 countries.

We’re looking for a one-time report that helps us understand our sales performance. As an entire company, we would like to see sales performance across time, broken down by year. We’re also interested in understanding how our sales vary between global regions. Finally, we’d like to know which of our products drive the biggest profit for the company.

## Available Data
Data are held on a spreadsheet, saved in csv format. The sheet contains a list of previous items sold.

LINK TO DATA SOURCE: ORDERS

## Deliverables
We want a one-time report that includes the following items:

* Summarize the company’s yearly sales performance, including:    
  + New customers
  + Total number of orders placed
  + Total number of items sold
  + Total sales volume ($)

* We store the world region of each sale. We’d like to know the following broken down by sales region:
  + Number of orders placed
  + Number of items sold
  + Total sales volume ($)

* Come up with a method to discover which of our products makes the company the most profit.

---

```{r include=FALSE, results='hide', message=FALSE, warning=FALSE}
# LOAD LIBRARIES
library(tidyverse)

# PREP WORKSPACE
rm(list = ls())
all_data <- read.csv('C:/Users/danie/Documents/R Scripts/analytics_portfolio/case_study_global_superstore/data/global_superstore_2016_orders.csv')

# REDUCE DATA 
all_data <- all_data %>% 
  rename(OrderId = Order.ID,
         RowId = Row.ID,
         OrderDate = Order.Date,
         ShipDate = Ship.Date,
         ShipMode = Ship.Mode,
         CustomerId = Customer.ID,
         CustomerName = Customer.Name,
         PostalCode = Postal.Code,
         ProductId = Product.ID,
         SubCategory = Sub.Category,
         ProductName = Product.Name,
         ShippingCost = Shipping.Cost,
         OrderPriority = Order.Priority) %>% 
  select(OrderId, Region, OrderDate, CustomerId, Segment,
         Market, ProductId, Category, SubCategory, ProductName, 
         Sales, Quantity, Profit)

# FORMAT DATES
all_data$OrderDate <- as.Date(all_data$OrderDate, '%m/%d/%Y')

# GET ORDER YEAR
all_data$OrderYear <- format(all_data$OrderDate, '%Y')

# FORMAT DOLLAR AMOUNTS
all_data$Sales = gsub("\\$", "", all_data$Sales)
all_data$Sales = gsub("\\,", "", all_data$Sales)
all_data$Sales = as.numeric(all_data$Sales)

all_data$Profit = gsub("\\$", "", all_data$Profit)
all_data$Profit = gsub("\\,", "", all_data$Profit)
all_data$Profit = as.numeric(all_data$Profit)

# FIND NEW CUSTOMERS 
new_customers <- all_data %>% 
  group_by(CustomerId) %>% 
  summarise(FirstOrder = min(OrderDate)) %>% 
  ungroup() %>% 
  mutate(cohort = format(FirstOrder, '%Y')) %>% 
  group_by(cohort) %>% 
  summarise(NewUsers = n()) %>% 
  ungroup()

# FIND YEARLY SALES
yearly_sales <- all_data %>% 
  group_by(OrderYear) %>% 
  summarise(OrdersFilled = n(), 
            ItemsSold = sum(Quantity),
            VolumeSold = sum(Sales))

# AGGREGATE REGIONAL SALES
regional_data <- all_data %>% 
  group_by(Region, OrderYear) %>% 
  summarise(OrdersFilled = n(), 
            ItemsSold = sum(Quantity),
            VolumeSold = sum(Sales)) %>% 
  ungroup()

```


# Yearly Sales Performance

### New customers by year:
```{r, echo=FALSE}
ggplot(new_customers, aes(x = cohort, 
                           y = NewUsers)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = NewUsers), 
              nudge_y = 200) +
    labs(title = 'New Customers by Year',
         x = '',
         y = 'New Customers') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```


### Number of orders placed by year:
```{r, echo=FALSE}
ggplot(yearly_sales, aes(x = OrderYear, 
                           y = OrdersFilled)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = OrdersFilled), 
              nudge_y = 1000) +
    labs(title = 'Total Order Count by Year',
         x = '',
         y = 'Orders Placed') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```

### Number of items sold by year:
```{r, echo=FALSE}
ggplot(yearly_sales, aes(x = OrderYear, 
                           y = ItemsSold)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = ItemsSold), 
              nudge_y = 2000) +
    labs(title = 'Total Items Sold by Year',
         x = '',
         y = 'Items Sold') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```

### Total sales revenue ($) by year:
```{r, echo=FALSE}
options(scipen = 10)
ggplot(yearly_sales, aes(x = OrderYear, 
                           y = VolumeSold)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = paste0('$', VolumeSold)), 
              nudge_y = 100000) +
    labs(title = 'Total Revenue ($) from Sales by Year',
         x = '',
         y = 'Sales Revenue ($)') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text.x = element_text(size = 12, color = 'black'),
          axis.text.y = element_text(size = 12, color = 'black', angle = 30))
```

---

# Sales Performance by Region
To get a better idea of how well each region has performed over time, I will group by ```Region``` and ```OrderYear``` and calculate performance metrics:

--- 

# Most Profitable Products