---
title: 'Case Study: A Sales Report for Global Superstore'
author: "Daniel Kuhman"
date: "`r Sys.Date()`"
output: html_document
---

### Business Prompt
Welcome to Global Superstore. We’re an online retailer specializing in office items, including basic supplies, furniture, and tech products. We serve corporate, consumer, and home-based buyers. As our name suggests, we operate globally, with customers from 165 countries.

We’re looking for a simple, one-time report that helps us understand our sales performance. As an entire company, we would like to see sales performance across time, broken down by year. We’re also interested in understanding how our sales varied between global regions in our most recent year of sales (2015) relative to the prior year. Specifically, we want to know how sales changed from 2014 to 2015 in each region. Finally, we’d like to know which of our products drive the biggest profit for the company.

### Available Data
Data are held in our BigQuery data warehouse in the ```orders``` table.


### Deliverables
We want a one-time report that includes the following items:

* Summarize the company’s yearly performance, including:    
  + Unique customers
  + Total number of orders placed
  + Total number of items sold
  + Total sales volume ($)

* We store the world region of each sale. We’d like to know the following broken down by sales region for 2015:
  + Percent change in number of orders placed from 2014 to 2015
  + Percent change in number of items sold from 2014 to 2015
  + Percent change in total sales volume ($) from 2014 to 2015

* Provide a list of the top 10 most profitable items

---

### Prepare R Workspace
To create visualizations for the data the client has requested, I will be using R. Prior to starting the analysis, I need to load libraries and clear the workspace in R:

```{r, include=TRUE, message=FALSE, warning=FALSE}
# LOAD LIBRARIES
library(tidyverse)
library(kableExtra)

# PREP WORKSPACE
rm(list = ls())

```

---

# Part 1: Summary of Yearly Sales Performance

<br>

### Extracting Data from BigQuery
The following SQL code will pull the data needed to assess yearly performance:
    
    WITH 

    allData AS (
      SELECT 
        Order_ID AS OrderId,
        EXTRACT(YEAR FROM Order_Date) AS OrderYear, --GET YEAR OF ORDER
        Customer_ID AS CustomerId, --WILL NEED TO FIND UNIQUE CUSTOMER IDS
        Quantity, --NUMBER OF PRODUCT SOLD IN THE ORDER
        Quantity * cast(Sales AS NUMERIC) AS TotalPrice
      FROM `global_superstore.orders` AS orders
    )

    SELECT 
      OrderYear,
      COUNT(DISTINCT CustomerId) AS UniqueCustomers,
      COUNT(DISTINCT OrderId) AS OrdersPlaced,
      SUM(Quantity) AS ItemsSold,
      SUM(TotalPrice) AS SalesRevenue
    FROM allData
    GROUP BY OrderYear
  
<br>

### Loading Data into R for Visualization

After running the SQL script above in BigQuery, I saved the output to a csv file. Here, I load the csv file into R before creating visualizations:

```{r, message=FALSE, warning=FALSE}
# LOAD THE DATA AND ASSIGN TO VARIABLE
yearly_data <- read.csv('C:/Users/danie/Documents/R Scripts/analytics_portfolio/case_study_global_superstore/data/yearly_sales.csv')

# VIEW THE DATA
yearly_data
```

<br>

### Data Visualizations
To show yearly performance in the metrics requested, I will use simple bar charts. To avoid repetition and clutter, I will provide code for the first chart only. The other charts will have the same format but with a different variable on the y-axis. 

```{r, fig.align='center'}
# BAR PLOT: UNIQUE CUSTOMERS BY YEAR
ggplot(yearly_data, aes(x = OrderYear,
                        y = UniqueCustomers)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = UniqueCustomers),
              nudge_y = 200) +
    labs(title = 'Unique Customers by Year',
         x = '',
         y = 'Unique Customers') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```

<br>

```{r, echo=FALSE, fig.align='center'}
# BAR PLOT: ORDERS BY YEAR
ggplot(yearly_data, aes(x = OrderYear,
                        y = OrdersPlaced)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = OrdersPlaced),
              nudge_y = 500) +
    labs(title = 'Total Orders Placed by Year',
         x = '',
         y = 'Orders Placed') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```

<br>


```{r, echo=FALSE, fig.align='center'}
# BAR PLOT: ITEMS SOLD BY YEAR
ggplot(yearly_data, aes(x = OrderYear,
                        y = ItemsSold)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = ItemsSold),
              nudge_y = 2000) +
    labs(title = 'Total Items Sold by Year',
         x = '',
         y = 'Items Sold') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text = element_text(size = 12, color = 'black'))
```

<br>

```{r, echo=FALSE, fig.align='center'}
# BAR PLOT: REVENUE BY YEAR
options(scipen = 10)
ggplot(yearly_data, aes(x = OrderYear,
                        y = SalesRevenue)) +
    geom_bar(stat = 'identity',
             color = 'black',
             fill = '#3366CC') +
    geom_text(aes(label = paste('$', SalesRevenue)),
              nudge_y = 1000000) +
    labs(title = 'Total Revenue from Sales by Year',
         x = '',
         y = 'Revenue ($)') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text.x = element_text(size = 12, color = 'black'),
          axis.text.y = element_text(size = 12, color = 'black', angle = 30))
```

---

# Part 2: Regional 

<br>

### Extracting Data from BigQuery
The following SQL code will pull the data needed to assess Regional sales data:

    WITH 
    
    allData AS (
      SELECT 
        Order_ID AS OrderId,
        EXTRACT(YEAR FROM Order_Date) AS OrderYear, --GET YEAR OF ORDER
        Region, --GLOBAL REGION
        Quantity, --NUMBER OF PRODUCT SOLD IN THE ORDER
        Quantity * cast(Sales AS NUMERIC) AS TotalPrice
    FROM `global_superstore.orders` AS orders
    WHERE EXTRACT(YEAR FROM Order_Date) IN (2014, 2015)
    ),
    
    agData AS (
      SELECT 
      Region,
      OrderYear,
      COUNT(DISTINCT OrderId) as OrdersPlaced,
      SUM(Quantity) as ItemsSold,
      SUM(TotalPrice) as SalesRevenue
    FROM allData
    GROUP BY Region, OrderYear
    ORDER BY Region, OrderYear
    )
    
    SELECT * 
    FROM (
      SELECT 
      Region,
      ((OrdersPlaced - LAG(OrdersPlaced) OVER(PARTITION BY Region ORDER BY OrderYear)) / LAG(OrdersPlaced) OVER(PARTITION BY Region ORDER BY OrderYear)) * 100 as PercDiffOrdersPlaced,
      ((ItemsSold - LAG(ItemsSold) OVER(PARTITION BY Region ORDER BY OrderYear)) / LAG(ItemsSold) OVER(PARTITION BY Region ORDER BY OrderYear)) * 100 as PercDiffItemsSold,
      ((SalesRevenue - LAG(SalesRevenue) OVER(PARTITION BY Region ORDER BY OrderYear)) / LAG(SalesRevenue) OVER(PARTITION BY Region ORDER BY OrderYear)) * 100 as PercSalesRevenue
    FROM agData) innerData
    WHERE PercDiffOrdersPlaced IS NOT NULL

<br>

### Loading Data into R for Visualization

After running the SQL script above in BigQuery, I saved the output to a csv file. Here, I load the csv file into R before creating visualizations:

```{r, message=FALSE, warning=FALSE}
# LOAD THE DATA AND ASSIGN TO VARIABLE
regional_data <- read.csv('C:/Users/danie/Documents/R Scripts/analytics_portfolio/case_study_global_superstore/data/regional_data.csv')

# VIEW THE DATA
regional_data
```

<br>

### Data Visualizations
To show Regional performance from 2014 to 2015 in the metrics requested, I will use simple, ordered bar charts. Bars will be filled based on value, with positive year over year changes in green and negative changes in red. To avoid repetition and clutter, I will provide code for the first chart only. The other charts will have the same format but with a different variable on the y-axis. 

```{r, echo=TRUE, fig.align='center', fig.width=10}
# BAR PLOT: CHANGE IN ORDERS BY REGION
ggplot(regional_data, aes(x = reorder(Region, -PercDiffOrdersPlaced),
                          y = PercDiffOrdersPlaced,
                          fill =  PercDiffOrdersPlaced < 0)) +
    geom_bar(stat = 'identity',
             color = 'black') +
    scale_fill_manual(values = c("#006600", "firebrick")) + # THIS WILL FORMAT THE FILL BASED ON VALUE (POS / NEG)
    labs(title = 'Percent Change in Orders Placed from 2014 to 2015',
         x = '',
         y = 'Change in Orders (%)') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text.x = element_text(size = 12, color = 'black', angle = 45, hjust = 1),
          axis.text.y = element_text(size = 12, color = 'black'),
          legend.position = 'none')
```

<br>

```{r, echo=FALSE, fig.align='center', fig.width=10}
# BAR PLOT: CHANGE IN ITEMS SOLD BY REGION
ggplot(regional_data, aes(x = reorder(Region, -PercDiffItemsSold),
                          y = PercDiffItemsSold,
                          fill =  PercDiffItemsSold < 0)) +
    geom_bar(stat = 'identity',
             color = 'black') +
    scale_fill_manual(values = c("#006600", "firebrick")) +
    labs(title = 'Percent Change in Total Items Sold from 2014 to 2015',
         x = '',
         y = 'Change in Items Sold (%)') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text.x = element_text(size = 12, color = 'black', angle = 45, hjust = 1),
          axis.text.y = element_text(size = 12, color = 'black'),
          legend.position = 'none')
```

<br>

```{r, echo=FALSE, fig.align='center', fig.width=10}
# BAR PLOT: CHANGE IN REVENUE BY REGION
ggplot(regional_data, aes(x = reorder(Region, -PercSalesRevenue),
                          y = PercSalesRevenue,
                          fill =  PercSalesRevenue < 0)) +
    geom_bar(stat = 'identity',
             color = 'black') +
    scale_fill_manual(values = c("#006600", "firebrick")) +
    labs(title = 'Percent Change in Sales Revenue from 2014 to 2015',
         x = '',
         y = 'Change in Sales Revenue (%)') +
    theme_bw() +
    theme(plot.title = element_text(size = 18, face = 'bold', hjust = 0.5),
          axis.title.y = element_text(size = 16, face = 'bold', margin = margin(t = 0, r = 15, b = 0, l = 0)),
          axis.text.x = element_text(size = 12, color = 'black', angle = 45, hjust = 1),
          axis.text.y = element_text(size = 12, color = 'black'),
          legend.position = 'none')
```

---

# Part 3: Most Profitable Items

<br>

### Extracting Data from BigQuery
The following SQL code will pull the data needed to find the top 10 most items ranked by greatest total profit. Note that the ```Product_Name``` field in the original dataset is formatted as ```Product Name, Description``` - we're only interested in the product name, not the description, so I extract the name using ```REGEX_SUBSTR()```. I also pull the total number of each product sold and the average profit per unit.

    SELECT *
    FROM (
      SELECT 
        REGEXP_SUBSTR(Product_Name, r"^(.+?),") AS ProductName,
        COUNT(*) AS TotalSold,
        ROUND(AVG(Quantity * cast(Profit as numeric)),2) AS AvgProfit,
        SUM(Quantity * cast(Profit as numeric)) AS TotalProfit
      FROM `global_superstore.orders` AS orders
      GROUP BY REGEXP_SUBSTR(Product_Name, r"^(.+?),")) innerData
    ORDER BY TotalProfit DESC
    LIMIT 10

<br>

### Loading Data into R for Visualization

After running the SQL script above in BigQuery, I saved the output to a csv file. Here, I load the csv file into R before creating visualizations:

```{r}
profit_prods <- read.csv('C:/Users/danie/Documents/R Scripts/analytics_portfolio/case_study_global_superstore/data/profitable_products.csv')

profit_prods
```

<br>

### Data Visualizations

To show the most profitable products, I will use a simple data table. 

```{r}
kable(profit_prods,
      col.names = c('Product',
                    'Total Number Sold',
                    'Average Per-Unit Profit (USD)',
                    'Total Profit (USD)'),
      align = c('l', 'c', 'c', 'c'),
      caption = 'Products ranked by total profits generated, all-time') %>% 
  kable_styling()
```


---